---
title: binfer
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/nicksolomon/binfer.svg?branch=master)](https://travis-ci.org/nicksolomon/binfer)
[![codecov](https://codecov.io/gh/nicksolomon/binfer/branch/master/graph/badge.svg)](https://codecov.io/gh/nicksolomon/binfer)

This package aims to make doing Bayesian things a little easier by giving a clear set of steps that you can easily use to simulate draws from a posterior distribution defined by a specific likelihood and prior.

## The steps

1. `define` a likelihood
  - Input: A data frame, a formula with the response variable and the name of a function that is your likelihood.
  - Output: A data frame with the name of the function given at input as an attribute
2. `assume` a prior distribution
  - Input: A data frame and a function of a single variable
  - Ouput: A data frame with the name of the function added as an attribute
3. `draw` from the posterior distribution
  - Input: A data frame, a control parameter
  - Output: A data frame of draws from the posterior distribution
4. `diagnose` the chain:
  - Input: a data frame of draws from the posterior
  - Output: The same data frame and a bunch of diagnostic plots
5. `clean` the chain
  - Input: Draws from the posterior
  - Output: Burned-in and subsampled draws from the posterior
6. Construct your confidence interval or p-value
  - Use `dplyr` for this
  
## Examples

Let's check our simulations by comparing the results to an analytically tractable problem. We'll use count data that we assume follows a poisson distribution. The prior we choose is the conjugate prior, a gamma distribution. The posterior should be a gamma distribution as well. Let's see if that's what we get!
```{r}
library(binfer)
library(tidyverse)

set.seed(20180127)

# Load taxis dataset
data("taxis")

# Define the prior and the likelihood
my_lik <- function(data, theta) {if (theta > 0) dpois(data, lambda = theta) else 0}
my_prior <- function(theta) {dgamma(theta, rate = 10, shape = 10)} 

# Sample the data to reduce runtimes
taxis_small <- taxis %>%
  sample_frac(.1)

# Do the binfer steps
posterior <- taxis_small %>% 
  define(passenger_count ~ my_lik) %>% 
  assume(~ my_prior) %>% 
  draw(initial = mean(taxis$passenger_count), nbatch = 1e5, blen = 1, scale = .1) %>% 
  diagnose() %>% 
  clean(burnin = 1000, subsample = 15) %>% 
  diagnose()

# Calculate the analytical hyperparameters
hyper1 <- 10 + sum(taxis_small$passenger_count)
hyper2 <- 10 + length(taxis_small$passenger_count)

# Analytic mean estimator
hyper1/hyper2

# Our mean estimator
posterior %>% 
  summarize(mean = mean(chain))

# Plot the simulated density against the analytical density
ggplot(posterior, aes(chain)) + 
  geom_density() +
  stat_function(fun = dgamma, args = list(hyper1, hyper2), color = "red")
```


Estimate the standard deviation of a normal distribution with a gamma prior:
```{r example1}
library(binfer)
library(tidyverse)

my_lik <- function(data, theta) {if (theta > 0) dnorm(data, mean = mean(iris$Sepal.Width) , sd = theta) else 0}
my_prior <- function(theta) {dgamma(theta, shape = 10, rate = 20)}

posterior <- define(iris, Sepal.Width ~ my_lik) %>% 
  assume(prior = ~ my_prior) %>% 
  draw(initial = .43, nbatch = 1e5, blen = 1, scale = .05) %>% 
  diagnose() %>% 
  clean(burnin = 0, subsample = 20) %>% 
  diagnose()

posterior %>% summarise(mean = mean(chain),
                        sd = sd(chain),
                        lower = quantile(chain, .025),
                        upper = quantile(chain, .975))
```

Estimate the probability of success of a binomnial distribution with a beta prior:

```{r example2}
binom_test_data <- rbinom(50, prob = .1, size = 1) %>% 
  tibble(response = .)

binom_lik <- function(data, theta) {if(theta > 0 & theta < 1) dbinom(data, prob = theta, size = 1) else 0}
beta_prior <- function(theta) {dbeta(theta, 1, 2)}

posterior2 <- binom_test_data %>% 
  define(response ~ binom_lik) %>% 
  assume(~ beta_prior) %>% 
  draw(initial = .5, nbatch = 1e5, blen = 1, scale = .15) %>% 
  diagnose() %>% 
  clean(burnin = 1000, subsample = 30) %>% 
  diagnose()
posterior2 %>% 
  summarize(mean = mean(chain), 
            sd = sd(chain), 
            lower = quantile(chain, .025), 
            upper = quantile(chain, .975))

```
